---
title: "LinkedIn Post — Figure Audit"
subtitle: "Verifying every claim in the narrative against the source data"
date: today
format:
  html:
    self-contained: true
    toc: true
    toc-depth: 2
    code-fold: true
    output-file: audit-figures-report.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
library(dplyr)
library(scales)
library(ggplot2)
```

## The LinkedIn post claims

The chart and post make the following specific, checkable claims:

> **Only 4 of 47 African countries** meet the SDG maternal-mortality target.
>
> Maternal deaths per 100,000 live births (WHO, 2023). **Nigeria's rate (993)** is **25× Cabo Verde's (40)**. The SDG 3.1 target is **< 70**.

This report verifies each bolded claim against `data/africa_mmr.csv`.

```{r load}
df <- read.csv("../data/africa_mmr.csv")
SDG_TARGET <- 70

meet_sdg <- df %>% filter(mmr < SDG_TARGET) %>% arrange(mmr)
n_total  <- nrow(df)
n_meet   <- nrow(meet_sdg)
```

## 1. "47 African countries"

```{r claim-47}
cat("Countries in CSV:", n_total, "\n")
cat("Year range:", min(df$year), "-", max(df$year), "\n")
```

```{r verdict-47, results='asis'}
if (n_total == 47) {
  cat("::: {.callout-tip}\n## PASS\nCSV contains exactly 47 African countries.\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\nCSV contains", n_total, "countries, not 47.\n:::\n")
}
```

## 2. "Only 4 meet the SDG target (< 70)"

```{r claim-sdg}
knitr::kable(
  meet_sdg %>% select(country, country_code, year, mmr),
  col.names = c("Country", "Code", "Year", "MMR"),
  caption = paste("Countries with MMR <", SDG_TARGET)
)
```

```{r verdict-sdg, results='asis'}
if (n_meet == 4) {
  cat("::: {.callout-tip}\n## PASS\n",
      n_meet, "countries have MMR <", SDG_TARGET, ".\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\n",
      n_meet, "countries have MMR <", SDG_TARGET,
      "— claim says 4.\n:::\n")
}
```

## 3. "Nigeria's rate (993)"

```{r claim-nigeria}
ng <- df %>% filter(country_code == "NGA")
knitr::kable(
  ng %>% select(country, year, mmr, low, high),
  col.names = c("Country", "Year", "MMR", "Low", "High"),
  format.args = list(big.mark = ",")
)
ng_mmr <- ng$mmr
```

```{r verdict-nigeria, results='asis'}
if (round(ng_mmr) == 993) {
  cat("::: {.callout-tip}\n## PASS\nNigeria's MMR rounds to 993.\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\nNigeria's MMR is",
      round(ng_mmr, 1), "— does not round to 993.\n:::\n")
}
```

## 4. "Cabo Verde's (40)"

```{r claim-cabo}
cv <- df %>% filter(country_code == "CPV")
knitr::kable(
  cv %>% select(country, year, mmr, low, high),
  col.names = c("Country", "Year", "MMR", "Low", "High"),
  format.args = list(big.mark = ",")
)
cv_mmr <- cv$mmr
```

```{r verdict-cabo, results='asis'}
if (round(cv_mmr) == 40) {
  cat("::: {.callout-tip}\n## PASS\nCabo Verde's MMR rounds to 40.\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\nCabo Verde's MMR is",
      round(cv_mmr, 1), "— does not round to 40.\n:::\n")
}
```

## 5. "25× Cabo Verde's rate"

```{r claim-ratio}
ratio <- ng_mmr / cv_mmr
cat("Nigeria:    ", round(ng_mmr), "\n")
cat("Cabo Verde: ", round(cv_mmr), "\n")
cat("Exact ratio:", round(ratio, 2), "x\n")
```

```{r verdict-ratio, results='asis'}
if (ratio >= 24.5 & ratio < 25.5) {
  cat("::: {.callout-tip}\n## PASS\nRatio of", round(ratio, 2), "rounds to 25.\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\nRatio is",
      round(ratio, 2), "— does not round to 25.\n:::\n")
}
```

## 6. Highest-MMR country = Nigeria?

The post calls out Nigeria as the cautionary example. Verify it's actually the worst.

```{r claim-highest}
top5 <- df %>% slice_max(mmr, n = 5) %>% select(country, country_code, mmr)
knitr::kable(
  top5,
  col.names = c("Country", "Code", "MMR"),
  format.args = list(big.mark = ",")
)
highest <- top5$country[1]
```

```{r verdict-highest, results='asis'}
if (highest == "Nigeria") {
  cat("::: {.callout-tip}\n## PASS\nNigeria is the highest-MMR country.\n:::\n")
} else {
  cat("::: {.callout-warning}\n## CHECK\nHighest-MMR country is",
      highest, "— not Nigeria.\n:::\n")
}
```

## 7. Context: distribution of MMR across Africa

```{r mmr-distribution, fig.width=8, fig.height=4}
ggplot(df, aes(x = mmr)) +
  geom_histogram(binwidth = 50, fill = "#242852", color = "white") +
  geom_vline(xintercept = SDG_TARGET, linewidth = 0.8,
             linetype = "dashed", color = "#E24A3F") +
  annotate("text", x = SDG_TARGET, y = Inf,
           label = paste("SDG target <", SDG_TARGET),
           hjust = -0.05, vjust = 2, size = 3.5,
           color = "#E24A3F", fontface = "bold") +
  scale_x_continuous(breaks = seq(0, 1000, 100)) +
  labs(
    title = "Distribution of maternal mortality ratios across 47 African countries",
    subtitle = "Each bar = number of countries in that 50-unit MMR band; red line = SDG 3.1 target",
    x = "MMR (deaths per 100,000 live births)",
    y = "Number of countries"
  ) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank())
```

## Summary

```{r summary-table}
results <- data.frame(
  Claim = c(
    '"47 African countries"',
    '"Only 4 meet the SDG target (< 70)"',
    '"Nigeria 993"',
    '"Cabo Verde 40"',
    '"25× ratio"',
    'Nigeria = worst in Africa'
  ),
  Actual = c(
    n_total,
    paste(n_meet, "country(ies) with MMR <", SDG_TARGET),
    round(ng_mmr, 1),
    round(cv_mmr, 1),
    paste0(round(ratio, 2), "x"),
    highest
  ),
  Result = c(
    ifelse(n_total == 47, "PASS", "CHECK"),
    ifelse(n_meet == 4, "PASS", "CHECK"),
    ifelse(round(ng_mmr) == 993, "PASS", "CHECK"),
    ifelse(round(cv_mmr) == 40, "PASS", "CHECK"),
    ifelse(ratio >= 24.5 & ratio < 25.5, "PASS", "CHECK"),
    ifelse(highest == "Nigeria", "PASS", "CHECK")
  )
)

knitr::kable(results, align = c("l", "r", "c"))
```

---

*Report generated `r Sys.time()` from `data/africa_mmr.csv`.*
